## This script pulls plate reader data from the "Away From Home" series of Coculture experiments
## This script ultimately produces figure 3, and uses area under the growth curve to determine statistical differences in algal growth.

#### Libraries ####
```{r setup}

library(tidyverse)
library(lubridate)
library(DescTools)
library(broom)
library(ggpubr)
library(rstatix)
library(growthrates)
# for Foreach/doParallel
library(foreach)
library(doParallel)
numCores <- detectCores()
registerDoParallel(numCores-1)

#### Read in Data From Master ####
colors <- read.csv("../csv_files/colors.csv")|>
  mutate(speciesCode = hostShort)|>
  select(speciesCode, color)

tax <- read.csv("../csv_files/collection_tax_data.csv") |>
  select(asv, Isolate, mixed, Phylum:Genus)
tax_full <- read.csv("../csv_files/collection_tax_data.csv")

extract_dfs <- function(x, full_results) { 
  tryCatch(
    expr = {
      num <- x
      name <- names(full_results[num]) 
      df <- as.data.frame(full_results[[num]]$coefficients) |>
        mutate(sample = name) |>
        dplyr::select(sample, `Std. Error`, Estimate) |>
        rename(growthrate = Estimate,
               std_error = `Std. Error`) |>
        slice(2) }, 
    finally = {
      return(df)
    }
  )
}

## Grubbs Test checks for outliers and gives statistical basis for exclusion
## Used just before normalization
grubbs <- function(df){
  data_grubbs <- split(df, df$Isolate, df$new_host)
  for(i in 1:length(data_grubbs)){
    x <- data_grubbs[[i]]
    data_grubbs[[i]] <- x |>
      mutate(p_high = grubbs.test(x$asymptote, type = 10)[[3]],
             p_low = grubbs.test(x$asymptote, type = 10, opposite = T)[[3]],
             maxVal = max(x$asymptote),
             minVal = min(x$asymptote),
             outlier = ifelse(p_high <= 0.05 & maxVal == asymptote, T,
                              ifelse(p_low <= 0.05 & minVal == asymptote, T, F)))
    # |>
    #   filter(outlier == F)|>
    #   mutate(p_high = grubbs.test(x$asymptote, type = 10)[[3]],
    #          p_low = grubbs.test(x$asymptote, type = 10, opposite = T)[[3]],
    #          maxVal = max(x$asymptote),
    #          minVal = min(x$asymptote),
    #          outlier = ifelse(p_high <= 0.05 & maxVal == asymptote, T,
    #                           ifelse(p_low <= 0.05 & minVal == asymptote, T, F)))
  }
  df_final <- bind_rows(data_grubbs)|>
    dplyr::select(-p_high, -p_low, -maxVal, -minVal)
  return(df_final)
}
```


#### Plate Data #####

```{r data import}
afh_plate_list <- list.files(path = "../raw_data/AFH_data", pattern = "^Plate_AW(\\d\\d|\\d)_Day_(\\d\\d|\\d).csv",
                         full.names = TRUE, ignore.case = TRUE)

read_plate_afh <- function(afh_plate_list) {
  read_csv(afh_plate_list, show_col_types = F)|>
    mutate(plate_no = str_extract(str_extract(afh_plate_list,
                                              pattern = "Plate_AW(\\d\\d|\\d)"),
                                  pattern = regex("(\\d\\d|\\d)")),
           filename = afh_plate_list,
           read_day = str_extract(str_extract(afh_plate_list,
                                              pattern = "(\\d\\d|\\d).csv"),
                                  pattern = regex("(\\d\\d|\\d)")),
           read_time = as.POSIXct(`Reading Date/Time`, 
                                  tryFormats = c("%m/%d/%Y %H:%M",
                                                 "%m/%d/%Y %H:%M:%S")))
}

afh_plateDataAll <-
  afh_plate_list |> 
  map_df(~read_plate_afh(.))|>
  rename(ChlA_100 = "Mean RFU [ChlA_100:460,685]")|>
  select(Well, ChlA_100, read_time, read_day, plate_no)

#### Map Data ####
afh_map_list <- list.files(path = "../raw_data/AFH_data",
                       pattern = "^plate_AW(\\d\\d|\\d)_map.csv",
                       full.names = TRUE, ignore.case = TRUE)

read_map_afh <- function(afh_map_list) {
  read_csv(afh_map_list, col_types = cols(.default = col_character())) |>
    mutate(Sample = as.character(Sample),
           plate_no = str_extract(str_extract(afh_map_list, pattern = "plate_AW(\\d\\d|\\d)"),pattern = regex("(\\d\\d|\\d)")), 
           filename = afh_map_list) |>
    select(-filename)
}

afh_mapDataAll <- 
  afh_map_list |>
  map_df(~read_map_afh(.))


#### Key Files ####
isolate_key <- read_csv("../raw_data/AFH_data/plate_map_isolate_codes.csv")

species_key <- read_csv("../raw_data/AFH_data/plate_map_species_codes.csv")

#### Assemble Data ####
afhData <- inner_join(afh_plateDataAll, afh_mapDataAll,
                   by = c("Well", "plate_no")) |>
  mutate(Sample = ifelse(Sample == "MC", "MC_MC", Sample))|>
  separate(Sample, c("sampleCode", "speciesCode"), remove = F)|>
  left_join( isolate_key, by = "sampleCode") |>
  left_join( species_key, by = "speciesCode") |>
  mutate(plateCode = paste(plate_no, speciesCode, sep = "_"),
         collectionCode = ifelse(sampleCode == "AC", "AC", collectionCode),
         hostCode = ifelse(sampleCode == "AC", speciesCode, hostCode),
         sample_exact = paste("plate", plate_no, "well", Well, Sample, sep = "_"),
         afh = ifelse(sampleCode == "AC", "Axenic", 
                      ifelse(sampleCode == "MC", "Media",
                             ifelse(speciesCode == hostCode, "Native", "Non-Native"))))|>
  # calc time replacement
  group_by(sample_exact)|>
  mutate(begin = min(read_time),
         read_interval = begin %--% read_time,
         read_timeHours = as.numeric(as.duration((read_interval)/dhours(1))))|>
  select(-begin)|>
  group_by(plate_no, Sample, read_day)|>
  mutate(day_mean = mean(ChlA_100, na.rm=TRUE),
         day_sd = sd(ChlA_100, na.rm=TRUE))|>
  ungroup()|>
  mutate(read_day = as.numeric(read_day))|>
  filter(read_day <= 12) %>%
  arrange(read_day)
```
```{r Growthrate and Carrying Capacity Calculations}
#### Calc Growthrate ####
gr_chlA <- afhData |>
  filter(collectionCode != "MC") |>
  all_easylinear(ChlA_100 ~ read_timeHours | sample_exact)
results_full_chlA <- summary(gr_chlA) %>%
  discard(~ length(.x) == 3)

results_chlA<- results(gr_chlA) |>
  rename(growthrate = mumax)

std_err_num <- as.list(1:length(results_full_chlA)) 

std_err_chlA <- lapply(std_err_num, extract_dfs,
                       full_results = results_full_chlA) |>
  bind_rows() |>
  mutate(yplus = growthrate + std_error,
         yminus = growthrate - std_error) |>
  dplyr::select(-growthrate)

#Add standard error to the dataframe
growthrates <- inner_join(results_chlA, std_err_chlA, by = c("sample_exact" = "sample")) |>
  dplyr::select(sample_exact, growthrate,
                r2, std_error) |>
  mutate(growthrate = ifelse(is.na(growthrate), 0, growthrate))

afhData_split <- split(afhData, afhData$sample_exact)

coefList <- foreach(i = 1:length(afhData_split), .packages = "tidyverse") %dopar%{
  tryCatch({
    df <- afhData_split[[i]]
    aPara <- max(df$ChlA_100)+ 0.05*max(df$ChlA_100)
    parameters <- coef(lm(qlogis(ChlA_100/aPara) ~ read_timeHours, data = df))
    iPara <- parameters[[1]]
    xPara <- parameters[[2]]
    
    mod <- nls(ChlA_100 ~ A/(1 + exp(-(I + X * read_timeHours))),
               start = list(A = aPara, I = iPara, X = xPara),
               data = df, trace = T, nls.control(maxiter = 100, warnOnly = T))
    modSummary <- summary(mod)
    
    aMod <- coef(mod)[1]
    iMod <- coef(mod)[2]
    xMod <- coef(mod)[3]
    
    coefDF <- data.frame(sample_exact = unique(df$sample_exact),
                         asymptote = aMod,
                         intercept = iMod,
                         growthParam = xMod)
    return(coefDF)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

iKey <- afhData |>
  dplyr::select(sample_exact, collectionCode, speciesCode, hostCode, plate_no, afh)|>
  distinct()

coefs <- bind_rows(coefList)|>
  left_join(iKey)|>
  left_join(growthrates) |>
  mutate(growthrate = ifelse(is.na(growthrate), growthParam, growthrate))

ACcoefs <- coefs |> filter(collectionCode == "AC")

coefList_4t <- split(coefs, list(coefs$collectionCode, coefs$speciesCode, coefs$plate_no), drop = T)#remove empty lists from resulting split dataframe

statsList <- foreach(i = 1:length(coefList_4t), .packages = c("tidyverse", "broom")) %dopar%{
  tryCatch({
    tt_df <- coefList_4t[[i]]
    new_host <- unique(tt_df$speciesCode)
    orig_host <- unique(tt_df$hostCode)
    plate <- unique(tt_df$plate_no)
    Isolate <- unique(tt_df$collectionCode)
    afh <- unique(tt_df$afh)
    ac_df <- ACcoefs |> filter(hostCode == new_host & plate_no == plate)
    cc_greater <- tidy(t.test(tt_df$asymptote, 
                              ac_df$asymptote, 
                              alternative = "greater"))|>
      mutate(pCC_greater = p.value, Isolate = Isolate)|>
      dplyr::select(pCC_greater, Isolate)
    cc_less <- tidy(t.test(tt_df$asymptote, 
                           ac_df$asymptote, 
                           alternative = "less"))|>
      mutate(pCC_less = p.value, Isolate = Isolate)|>
      dplyr::select(pCC_less, Isolate)
    gr_greater <- tidy(t.test(tt_df$growthrate, 
                              ac_df$growthrate, 
                              alternative = "greater"))|>
      mutate(pGR_greater = p.value, Isolate = Isolate)|>
      dplyr::select(pGR_greater, Isolate)
    gr_less <- tidy(t.test(tt_df$growthrate, 
                           ac_df$growthrate, 
                           alternative = "less"))|>
      mutate(pGR_less = p.value, Isolate = Isolate)|>
      dplyr::select(pGR_less, Isolate)
    cc_stat <- full_join(cc_greater, cc_less)
    gr_stat <- full_join(gr_greater, gr_less)
    stats_df <- full_join(cc_stat, gr_stat)|>
      mutate(plate_no = plate,
             new_host = new_host,
             orig_host = orig_host,
             afh = afh)
    return(stats_df)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

stats <- bind_rows(statsList)|>
  dplyr::select(Isolate, new_host, orig_host, plate_no, pCC_greater, pCC_less, pGR_greater, pGR_less)|>
  mutate(ccEffect = ifelse(pCC_greater <= 0.05 & pCC_less > 0.05, "Increased CC",
                           ifelse(pCC_greater > 0.05 & pCC_less <= 0.05, "Decreased CC",
                                  ifelse(pCC_greater > 0.05 & pCC_less > 0.05,
                                         "No Significant CC Change", "Error"))),
         grEffect = ifelse(pGR_greater <= 0.05 & pGR_less > 0.05, "Increased GR",
                           ifelse(pGR_greater > 0.05 & pGR_less <= 0.05, "Decreased GR",
                                  ifelse(pGR_greater > 0.05 & pGR_less > 0.05,
                                         "No Significant GR Change", "Error"))),
         annotation = paste(ccEffect, grEffect, sep = "\n"))

library(outliers)
stats_coefs <- left_join(stats, coefs, by = c("Isolate" = "collectionCode", 
                                              "new_host" = "speciesCode",
                                              "orig_host" = "hostCode",
                                              "plate_no"))|>
  grubbs() # This test added here when normalization was added 01/19/2022 -JL

```

#### Normalization of Plate Reader Data to Axenic Controls ####

```{r Data Normalization}
ac_coefs <- stats_coefs |> 
  filter(Isolate == "AC" & outlier != T)|>
  group_by(new_host, plate_no)|>
  mutate(nSamples = n(),
         meanCC = mean(asymptote),
         meanGR = mean(growthrate))|>
  ungroup()|>
  # These columns are empty, as there is no statistics data for axenic controls samples
  dplyr::select(-pCC_greater, -pCC_less, -pGR_greater, -pGR_less,
                -ccEffect, -grEffect, -annotation)

stats_coefs_split <- stats_coefs |>
  filter(Isolate != "AC" & Isolate != "MC")|>
  mutate(split_var = paste(Isolate, plate_no, sep = "_"))%>%
  split(., .$split_var)

for(i in 1:length(stats_coefs_split)){
  df <- stats_coefs_split[[i]]
  plate <- unique(df$plate_no)
  host <- unique(df$new_host)
  ac <- ac_coefs |>
    filter(new_host == host & plate_no == plate & asymptote >= 0)|>
    summarize(acCC = mean(asymptote),
              acGR = mean(growthrate))
  acCC <- as.numeric(ac$acCC)
  acGR <- as.numeric(ac$acGR)
  stats_coefs_split[[i]] <- df |>
    mutate(normCC = asymptote/acCC,
           normGR = growthrate/acGR,
           logNormCC = log(normCC),
           logNormGR = log(normGR))
}
stats_normCoefs <- bind_rows(stats_coefs_split)

#### Calculate Triplicate Mean of Stats Coefficients ####
stats_meanCoefs <- stats_normCoefs |>
  group_by(Isolate, new_host, plate_no) |>
  mutate(n = n(),
         meanCC = mean(asymptote),
         seCC = sd(asymptote) / sqrt(n),
         meanGR = mean(growthrate),
         seGR = sd(growthrate) / sqrt(n),
         mean_normCC = mean(normCC),
         mean_normGR = mean(normGR),
         mean_logNormCC = mean(logNormCC),
         mean_logNormGR = mean(logNormGR)
  ) |>
  filter(n >= 3) |>
  distinct(Isolate, plate_no, .keep_all = T)
```
#Calculate area under the curve for a timeseries for each well on each plate.
``` {r Area Under Curve Calculation}
aucData_split <- as.list(1:length(afhData_split)) 

for(i in 1:length(afhData_split)){
  df <- afhData_split[[i]]
  auc <- AUC(x = df$read_time, y = df$ChlA_100, method = "spline")
  aucData_split[[i]] <- data.frame("sample_exact" = df$sample_exact,
                              "Sample" = df$Sample,
                              "sampleCode" = df$sampleCode,
                              "plateCode" = df$plateCode,
                              "auc" = auc,
                              "Isolate" = df$collectionCode,
                              "orig_host" = df$hostCode,
                              "new_host" = df$speciesCode)|>
    distinct()
}

#Combine area under the curve data into one list.
aucData <- bind_rows(aucData_split)

ctrl_aucData <- filter(aucData, sampleCode == "AC")
sample_aucData <- filter(aucData, sampleCode != "AC")

#no outliers, not needed in code
# acAUC_outliers <- ctrl_aucData |>
#   group_by(plateCode) |>
#   identify_outliers(auc)
#ctrl_aucData <- anti_join(ctrl_aucData, acAUC_outliers, by = "auc")

sample_aucData_split <- aucData |>
  filter(Isolate != "AC" & Isolate != "MC") %>%
  split(list(.$plateCode, .$sampleCode), drop =  T)

#T-test between coculture and axenic controls, by plate
tTestData_split <- foreach(i = 1:length(sample_aucData_split), .packages = c("tidyverse","broom")) %dopar% {
  tryCatch({
  df <- sample_aucData_split[[i]]
  plate <- unique(df$plateCode)
  sample <- unique(df$Sample)
  
  ctrl <- filter(ctrl_aucData, plateCode == plate)
  
  stat_greater <- tidy(t.test(df$auc, ctrl$auc, alternative = "greater"))|>
    mutate(Sample = sample,
           p_greater = p.value) |>
    select(Sample, p_greater)
  
  stat_less <- tidy(t.test(df$auc, ctrl$auc, alternative = "less"))|>
    mutate(Sample = sample,
           p_less = p.value) |>
    select(Sample, p_less)
  
  tTestData <- full_join(stat_greater, stat_less)|>
    mutate(Effect = case_when((p_greater > 0.05 & p_less > 0.05) ~ "Not Significant",
                              p_greater <= 0.05 ~ "Positive",
                              p_less <= 0.05 ~ "Negative",
                              T ~ "Error"))
  return(tTestData)
  
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
}

tTestData <- bind_rows(tTestData_split)

#### AUC Normalization ####
AC_auc_mean <- ctrl_aucData |>
  group_by(plateCode) |>
  mutate(acAuc_mean = mean(auc),
         acAuc_se = sd(auc)/sqrt(n())) |>
  select(new_host, plateCode, acAuc_mean, acAuc_se) |>
  distinct()

auc_data_norm <- sample_aucData |>
  filter(! Isolate == "MC") |>
  left_join(AC_auc_mean, by = c("plateCode", "new_host"))|>
  mutate(auc_norm = auc/acAuc_mean,
         log_auc_norm = log(auc_norm)) |>
  left_join(tTestData) |>
  filter(!is.na(Effect))

```

```{r Map to Taxonomy Data}
#Rename to current naming conventions to match up to taxonomy data
sam.dataMiseqNames <- auc_data_norm |>
  mutate(Isolate = str_replace(Isolate, ",|\\.", "point"),
         Isolate = str_replace(Isolate, "DF", "D31"),
         seq_type = "Miseq") 

sam.dataJinny <- auc_data_norm |> 
  filter(str_detect(Isolate, "DF") & orig_host == "CS") |>
  #Match up experimental Isolates to the ones Jinny Sequenced that are missing
  mutate(Isolate = case_when(Isolate == "10DF" ~ "S10OMO",
                             Isolate == "11DF" ~ "S11",
                             Isolate == "12DF" ~ "S12",
                             Isolate == "13DF" ~ "S13",
                             Isolate == "14DF" ~ "S14",
                             Isolate == "16DF" ~ "S16",
                             Isolate == "17DF" ~ "S17A",
                             Isolate == "18DF" ~ "S18",
                             Isolate == "19DF" ~ "S19AB",
                             Isolate == "2DF" ~ "S2",
                             Isolate == "20DF" ~ "S20",
                             Isolate == "21DF" ~ "S21",
                             Isolate == "22DF" ~ "S22",
                             Isolate == "24DF" ~ "S24",
                             Isolate == "25DF" ~ "S25",
                             Isolate == "26DF" ~ "S26",
                             Isolate == "27DF" ~ "S27",
                             Isolate == "28DF" ~ "S28",
                             Isolate == "29DF" ~ "S29",
                             Isolate == "3DF" ~ "S3",
                             Isolate == "30,1DF" ~ "S30W",
                             Isolate == "31DF" ~ "S31",
                             Isolate == "32DF" ~ "S32AB",
                             Isolate == "33DF" ~ "S33",
                             Isolate == "34DF" ~ "S34",
                             Isolate == "31DF" ~ "S31",
                             Isolate == "5DF" ~ "S5",
                             Isolate == "6DF" ~ "S6",
                             Isolate == "7DF" ~ "S7",
                             Isolate == "8DF" ~ "S8",
                             Isolate == "9DF" ~ "S9W"
  ),
  seq_type = "Miseq") |>
  filter(!is.na(Isolate))

sam.dataSangerNames <- auc_data_norm |>
  mutate(seq_type = "Sanger",
         Isolate = str_replace(Isolate, "D3", "_D3"),
         Isolate = str_replace(Isolate, ",|\\.", "point"),
         Isolate = str_replace(Isolate, "DF", "_DF"),
         Isolate = str_replace(Isolate, "D31", "_D31"))

sam.dataDF <- rbind(sam.dataMiseqNames, sam.dataSangerNames, sam.dataJinny)

sam.data <- sam.dataDF |>
  filter(! Isolate == "AC") |>
  group_by(sample_exact) |>
  distinct(Isolate, .keep_all = T) |>
  mutate(Isolate = case_when(Isolate == "8aD3" ~ "8D3",
                             T ~ Isolate))

renamed_data <- sam.data |> 
  left_join(tax) |>
  filter(!is.na(asv), mixed == F) |>
  distinct(sample_exact, .keep_all = T) |>
  mutate(host_status = case_when(orig_host == new_host ~ "Original",
                                 T ~ "Transplant")) |>
  group_by(plateCode, sampleCode) |>
  mutate(mean_log_auc_norm = mean(log_auc_norm),
            se_log_auc_norm = sd(log_auc_norm)/sqrt(n())) |>
  distinct(sampleCode, plateCode, .keep_all =  T) |>
  ungroup()

orig_host_data <- renamed_data |> 
  filter(host_status == "Original") |>
  pivot_wider(names_from = c(host_status), values_from = c(mean_log_auc_norm, se_log_auc_norm)) |>
  select(Isolate, orig_host, mean_log_auc_norm_Original, se_log_auc_norm_Original, Effect) |>
  rename("recent_effect" = Effect)

orig_host_comp <- filter(tax_full, Isolate %in% orig_host_data$Isolate) |>
  select(Isolate, host_species, grEffect, ccEffect, Effect, mixed, seq_type) |>
  left_join(orig_host_data) 
  
  
comparison_data <- renamed_data |>
  filter(host_status == "Transplant") |>
  pivot_wider(names_from = c(host_status), values_from = c(mean_log_auc_norm, se_log_auc_norm)) |>
  left_join(orig_host_data, by = c('Isolate','orig_host')) |>
  select(Isolate, new_host, orig_host, Effect, asv:se_log_auc_norm_Original)

lm <- lm(mean_log_auc_norm_Transplant ~ mean_log_auc_norm_Original:orig_host, data = comparison_data)
summary(lm)

```


#### Plots ####

```{r Plots, }
#Scatterplot ordination
scatterplot <- ggplot(comparison_data,
                        aes( x = mean_log_auc_norm_Original, 
                         y = mean_log_auc_norm_Transplant, color = Effect, shape = new_host)) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Not Significant" = "gray",
                                "Positive" = "#096C72", 
                                "Negative" = "#D2042D"), 
                     "Growth Outcome on Transplant Host") +
  scale_shape_manual(labels = c("CS" ="Chorella",
                                "CM" = "Coelastrum",
                               "SA"= "Scenedesmus",
                               "MM" = "Monoraphidium",
                                "SC" = "Selenastrum"),
                       values = c("CS" = 0,
                                  "CM" = 1,
                                  "SA" = 2,
                                  "MM" = 5,
                                  "SC" = 6),
                       name = "Transplant Host") +
  geom_hline(yintercept = 0, color = "grey", linetype = "longdash") +
  geom_vline(xintercept = 0, color = "grey", linetype = "longdash") +
  ylab("Log Norm AUC Transplant Host") +
  xlab("Log Norm AUC Original Host") +
  geom_errorbar(aes(ymin = mean_log_auc_norm_Transplant - se_log_auc_norm_Transplant ,
                    ymax = mean_log_auc_norm_Transplant + se_log_auc_norm_Transplant)) +
  geom_errorbarh(aes(xmin = mean_log_auc_norm_Original - se_log_auc_norm_Original,
                     xmax = mean_log_auc_norm_Original + se_log_auc_norm_Original)) +
  theme_pubclean() +
  theme(legend.position = "bottom",
        legend.justification = "center",
        legend.box = "vertical") +
  labs(title = "Growth Outcomes of Original vs Transplanted Host Algae")

png(filename = "cross_host_outcomes.png",
    res = 450,
    type = "cairo",
    units = "in",
    width = 8,
    height = 8)
scatterplot
dev.off()



```
